---
title: "Bayesian linear regression"
author: "Lauren James"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(blm)
```

## `blm` package description

This package constructs a Bayesian Linear Model, which is a statistical linear regression approach that uses Bayesian inferences. It has a similar structure and appearance as the `lm` and `glm` interfaces. The Bayesian Linear Model takes on the structure y ~ x (+ other predictors) and provides a training set from which to predict response variables from real data.

This package was built as part of the course "Data Science: Software Development and Testing" during 2016-17 at the Bioinformatics Research Centre (BiRC) at Aarhus University, and is still under development using GitHub. 

## Using the `blm` package

### Installation

As this package is still under development, it can only be accessed and installed through github. In order to install a package through github, it requires the `devtools` package in R. Ensure that it is installed:

```r
install.packages("devtools")
library(devtools)
```

Then install and load the package using the `install_github` feature:

```r
install_github("laurenejames/blm")
library(blm)
```

### Features of the `blm` package

There are three built-in functions that aid in the set up for the constructor in the the main function `blm()` which are not so important, but will be listed here for information purposes:

`make_prior()`

This creates a prior distribution, with the following inputs `make_prior(model, alpha, mu = c(0,0))` where the model is a formula in the `blm` format mentioned above.

`model_matrix_responseless()`

Required to give the phiX matrix for the posterior distribution, with the following inputs `model_matrix_responseless(model, ...)` where the `...` indicates an optional data frame addition.

`update()`

Updates the prior distribution using observed data to generate a posterior distribution, with the following inputs `update(model, prior, beta, ...)`

**The main function: the `blm` constructor**

`blm()`

This produces the Bayesian Linear Modelm and the object for the functions mentioned below. 

*Functions associated with `blm()`*

* `coef()`
Returns the coefficients of the blm(), mean and Sigma

* `confint()`
Returns a dataframe containing the mean, and confidence intervals (level set by an input to the function). Input parameters are as follows, `confint(object, parm, level, ...)` where the object is the blm, parm is the parameter(s) in question (default considers all parameters) and level sets the confidence interval (default is 95%).

* `deviance()`
Returns the deviance of the blm object when fitting a model and predicting the response variable based on the fitted blm. 

* `fitted()` 
Returns the fitted response variables from the blm 

* `plot()` 
Plots the residuals vs. the fitted response variable for the blm object.

* `predict()`
Predicts response variables using a new data set from the fitted blm object. If there is no new data inputed into the function, it predicts based on the fitted response variable.

* `print()` 
Prints the following features of the blm object:
 + Fitted blm 
 + The fitted model's coefficients

* `residuals()` 
Returns the residual values from the predicted response variable

* `summary()`
Returns the following features of the blm object:
 + Fitted blm 
 + The fitted model's coefficients
 + The fitted model's residuals
 + The fitted model's deviance


### Example of using the `blm` package

```{r}
# Create simulation data
alpha = 2; beta = 0.4
w0 = 0.5; w1 = 0.2; w2 = 1.2
simdata = data.frame(x = rnorm(500), z = rnorm(500))
simdata$y = rnorm(500, w0+w1*simdata$x+w2*simdata$z, 1/beta)
simmodel = y ~ x + z

# Construct the blm
my_new_blm = blm(simmodel, make_prior(simmodel, alpha, mu = c(0,0)), beta, simdata)

# Exploit all the functions
coef(my_new_blm)
confint(my_new_blm)
fitted(my_new_blm)

## Second data set for the predict function
simdata2 = data.frame(x = rnorm(500), z = rnorm(500))
simdata2$y = rnorm(500, w0+w1*simdata2$x+w2*simdata2$z, 1/beta)

# Carry on exploiting the functions
predict(my_new_blm, simdata2, beta)
residuals(my_new_blm)
deviance(my_new_blm)
plot(my_new_blm)
print(my_new_blm)
summary(my_new_blm)
```


## Reference

Mailund, T. "Introduction to Data Science and Statistical Programming in R"
